---
title: "Hansard Summary Statistics"
format: pdf
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

```{r}
# environment setup
library(tidyverse)
library(arrow)

# data import
corpus <- read_parquet("../inputs/corpus_1998_to_2025.parquet")

# filter for just chamber proceedings
corpus <- corpus %>% filter(fedchamb_flag==0)
```

## House of Reps Summary Statistics

```{r}
bind_rows(
  corpus %>% distinct(date) %>% count(name="Value") %>% 
    mutate(Metric="Number of sitting days"),
  corpus %>% distinct(uniqueID) %>% filter(!is.na(uniqueID)) %>% count(name="Value") %>% 
    mutate(Metric="Number of unique speakers"),
  corpus %>% distinct(partyAbbrev) %>% filter(!is.na(partyAbbrev)) %>% count(name="Value") %>% 
    mutate(Metric="Number of unique parties"),
  corpus %>% distinct(electorate) %>% filter(!is.na(electorate)) %>% count(name="Value") %>% 
    mutate(Metric="Number of unique electorates")) %>% 
  select(Metric, Value) %>% 
  knitr::kable()
```

## Number of rows flagged as interjections for House of Reps

```{r}
corpus %>% 
  group_by(interject) %>% 
  count() %>% 
  ungroup() %>% 
  rename("Interjection Flag" = interject) %>% 
  knitr::kable()
```

## Number and proportion of interjections by gender for House of Reps

```{r}
left_join(
  corpus %>% group_by(gender, interject) %>% 
  count() %>% ungroup() %>% filter(interject==1),
  corpus %>% group_by(gender, interject) %>% 
  count() %>% ungroup() %>% pivot_wider(names_from = interject, values_from = n) %>% 
  mutate(prop_interject = `1`/(`1`+`0`)*100) %>% 
  select(-`0`, -`1`) %>% 
  mutate(prop_interject=round(prop_interject,2),
         prop_interject = paste0(prop_interject,"%"))) %>% 
  select(Gender=gender, Count=n, Proportion=prop_interject) %>% 
  mutate(Gender=str_to_sentence(Gender)) %>% 
  knitr::kable()
```
## Number of interjections by political party

```{r}
corpus %>% 
  filter(interject==1 & !is.na(partyName)) %>% 
  group_by(partyName) %>% 
  count() %>% 
  arrange(desc(n)) %>% 
  rename(`Party Name`=partyName) %>% 
  knitr::kable()
```

## Number of interjections by political party and gender

```{r}
corpus %>% 
  filter(interject==1 & !is.na(partyName) & !is.na(gender)) %>% 
  group_by(partyName, gender) %>% 
  count() %>% 
  ungroup() %>% 
  pivot_wider(names_from = gender, values_from = n) %>% 
  rename(`Party Name`=partyName, Female=female, Male=male) %>% 
  mutate(across(c(Female, Male), ~ifelse(is.na(.x), 0, .x))) %>% 
  knitr::kable()
```

## Average, standard deviation, minimum and maximum speech length (by number of words)

```{r}
speech_length_stats <- corpus %>% 
  group_by(speech_no, date) %>% 
  # add 1 because counting number of spaces
  summarise(speech_length = sum(str_count(body, "\\s+")+1)) %>% 
  ungroup()

speech_length_stats %>% 
  summarise(mean_speech_length = mean(speech_length, na.rm=T),
            min_speech_length = min(speech_length, na.rm=T),
            max_speech_length = max(speech_length, na.rm=T),
            sd_speech_length = sd(speech_length, na.rm=T)) %>% 
  pivot_longer(c(mean_speech_length:last_col()), names_to = "Metric", values_to = "Value") %>% 
    mutate(Value = round(Value, 2),
         Metric = case_when(
           Metric=="min_speech_length" ~ "Minimum",
           Metric=="max_speech_length" ~ "Maximum",
           Metric=="mean_speech_length" ~ "Average",
           Metric=="sd_speech_length" ~ "Standard Deviation")) %>% 
  knitr::kable()
```

## Number of days, per year for House of Reps

```{r}
corpus %>% 
  distinct(date) %>% 
  mutate(year = str_extract(date, "^\\d{4}")) %>% 
  group_by(year) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(year = as.numeric(year)) %>% 
  ggplot(aes(x=year, y=n)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x="Year", y="Number of sitting days",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of unique speakers per day

```{r}
corpus %>% 
  filter(!is.na(uniqueID)) %>% 
  distinct(date, uniqueID) %>% 
  group_by(date) %>% 
  count() %>%
  ggplot(aes(x=date, y=n)) +
  geom_point(alpha=0.5) +
  theme_bw() +
  theme(legend.position = "none") +
  labs(x="Year", y="Number of unique speakers",
     #  title = "Number of unique speakers per day",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of unique women and men speakers, per year

```{r}
corpus %>% 
  mutate(year = as.numeric(str_extract(date, "^\\d{4}"))) %>% 
  distinct(uniqueID, gender, year) %>% 
  filter(!is.na(gender)) %>% 
  mutate(gender = case_when(gender=="female"~"Female",
                       gender=="male"~"Male")) %>% 
  group_by(year, gender) %>% 
  count() %>% 
  ggplot(aes(x=year, y=n, color = gender)) +
  geom_line() +
  geom_point() +
  theme_bw() +
  labs(x="Year", y="Number of unique speakers", color="Gender",
      # title = "Number of unique speakers by gender per year",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of unique speeches, per day for House of Reps

```{r}
corpus %>% 
  distinct(date, speech_no) %>% 
  group_by(date) %>% 
  count() %>% 
  ungroup() %>% 
  ggplot(aes(x=date, y=n))+
  geom_point(alpha=0.5) +
  theme_bw()+
  theme(legend.position = "none") +
  labs(x="Date", y="Number of speeches",
      # title = "Number of speeches per day",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of interjections by year for House of Reps

```{r}
corpus %>% 
  select(date, interject) %>% 
  mutate(year = as.numeric(str_extract(date, "^\\d{4}"))) %>% 
  group_by(year, interject) %>% 
  count() %>% 
  ungroup() %>% 
  filter(interject==1) %>% 
  ggplot(aes(x=year, y=n)) +
  geom_col() +
  theme_bw() +
  theme(legend.position = "none")+
  labs(x="Year", y="Number of interjections",
      # title = "Number of interjections per year",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of interjections by day for House of Reps

```{r}
corpus %>% 
  select(date, interject) %>% 
  group_by(date, interject) %>% 
  count() %>% 
  ungroup() %>% 
  filter(interject==1) %>% 
  ggplot(aes(x=date, y=n)) +
  geom_point(alpha=0.5) +
  theme_bw() +
  theme(legend.position = "none")+
  labs(x="Date", y="Number of interjections",
      # title = "Number of interjections per day",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```

## Number of interjections by day, by gender for House of Reps

```{r}
corpus %>% 
  select(date, gender, interject) %>% 
  mutate(gender = case_when(gender=="female"~"Female",
                       gender=="male"~"Male")) %>% 
  group_by(date, interject, gender) %>% 
  count() %>% 
  ungroup() %>% 
  filter(interject==1, !is.na(gender)) %>% 
  ggplot(aes(x=date, y=n, color=gender)) +
  geom_point(alpha=0.5) +
  theme_bw() +
  theme(legend.position = "none")+
  labs(x="Date", y="Number of interjections",
       #title = "Number of interjections per day",
       caption = "Note: Proceedings captured are from March 1998 to July 2025.")
```
